---
title: "Calibration of the Ocean Model"
author: "Pierre Barbillon"
date: "January 16, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
```

```{r packages,message=FALSE}
rm(list=ls())
library(ggplot2)
library(colorRamps)
library(gridExtra)
library(DiceKriging)
library(DiceDesign)
library(hetGP)
library(MCMCpack)
library(lhs)
library(mvtnorm)
library(parallel)
```


Loading simulator and GPEs (homGP, hetGP and seqhetGP which is hetGP with a sequential design)
```{r simandGPEs}
# functions
source("SimulatorAndFunctions.R")
source("FunctionCalibration.R")
```

Fixing seed
```{r seed}
seed=1234
set.seed(seed)
```



## Retrieving field data

```{r field}
Field = read.table("FieldData.csv",sep=" ")
head(Field)
Xfield = as.matrix(Field[,1:4])
Yfieldnoise = Field[,5]
n = nrow(Xfield)
Ym <- mean(Yfieldnoise)
Yv <- var(Yfieldnoise)
YfieldnoiseN <- (Yfieldnoise-Ym)/sqrt(Yv)
vareps = 2^2 
```

## Emulators

Either
```{r emulator load,eval=FALSE}
#load("EmulatorCalib4D.Rdata")
```

Or
```{r rebuilding emulator}
## From Static design
Static = read.table("staticDOEcalibration.csv")
Xsim = as.matrix(Static[,1:4])
Ysim = Static[,5]
YsimN = (Ysim-Ym)/sqrt(Yv)  
  
covtype <- "Matern5_2"
noiseControl <- list(g_min=1e-6, g_bounds=c(1e-6, 1), lowerDelta=log(1e-6))
settings <- list(linkThetas="none", initStrategy="smoothed", return.hom=TRUE)
lower <- c(0.01, 0.01, 0.001, 0.001) 
upper <- c(30, 30, 100, 100) 

het <- mleHetGP(Xsim, YsimN, lower=lower, upper=upper, covtype=covtype, noiseControl=noiseControl, 
                settings=settings, maxit=10000)
hom <- het$modHom

## Sequential design
Seq = read.table("seqDOEcalibration.csv")
Xseq = as.matrix(Seq[,1:4])
Yseq = Seq[,5]
YseqN = (Yseq-Ym)/sqrt(Yv)  
  
covtype <- "Matern5_2"
noiseControl <- list(g_min=1e-6, g_bounds=c(1e-6, 1), lowerDelta=log(1e-6))
settings <- list(linkThetas="none", initStrategy="smoothed", return.hom=TRUE)
lower <- c(0.01, 0.01, 0.001, 0.001) 
upper <- c(30, 30, 100, 100) 

seqhet <- mleHetGP(Xseq, YseqN, lower=lower, upper=upper, covtype=covtype, noiseControl=noiseControl, 
                settings=settings, maxit=10000)
```



## Calibration



Settings for MCMC
```{r MCMCsettings}
nMCMC = 1e6  #2e4
nburnin =  2e4 #1000
Vrw2 <- diag(c(2e-3, 2e-3, 5e-5,3e-5,1e-5,1e-5)) # variance of Random Walk
Uprior =c(.5,.5)
Thin = 100
```

Calibration with homGP
```{r homGPcalib}
homMLEdisc = RangeEstim(Xfield[,1:2],YfieldnoiseN,hom,Uprior,vareps)
calHom2 <- MCMCmetrop1R(likcalibrationwithdisc, XF=Xfield[,1:2], yF=YfieldnoiseN, GP=hom, Sigdisc=NULL
                       ,logfun=TRUE, theta.init=c(0.5,0.5,1,homMLEdisc$sigb,homMLEdisc$psi), 
                       burnin=nburnin, mcmc=nMCMC,V=Vrw2/2,thin=Thin)
dfcalhom = as.data.frame(calHom2)
names(dfcalhom) = c("u1","u2","s2f","s2b")
dfcalhom[,1:2] = dfcalhom[,1:2] * 900 +100
dfcalhom[,3] = dfcalhom[,3] * Yv
dfcalhom[,4] = dfcalhom[,4] * Yv
hom1 = ggplot(dfcalhom,aes(x=u1,stat(density)))+ xlab(expression(K[x])) + geom_density() + theme_bw() + xlim(50,1000) + geom_vline(aes(xintercept=700),color="red")
hom2 = ggplot(dfcalhom,aes(x=u2,stat(density))) + xlab(expression(K[y])) + geom_density() + theme_bw() +xlim(50,1000)  + geom_vline(aes(xintercept=200),color="red")
hom3 = ggplot(dfcalhom,aes(x=s2f,stat(density)))+ xlab(expression(sigma[epsilon]^2))  + geom_density() + theme_bw() + xlim(1,7) + geom_vline(aes(xintercept=4),color="red")
hom4 = ggplot(dfcalhom,aes(x=s2b,stat(density)))+ xlab(expression(sigma[MD]^2)) + geom_density() + theme_bw() + xlim(1,7) 
```



Calibration with hetGP
```{r hetGPcalib}
hetMLEdisc = RangeEstim(Xfield[,1:2],YfieldnoiseN,het,Uprior,vareps)
calhet2 <- MCMCmetrop1R(likcalibrationwithdisc, XF=Xfield[,1:2], yF=YfieldnoiseN, GP=het, Sigdisc=NULL
                        ,logfun=TRUE, theta.init=c(0.5,0.5,1,hetMLEdisc$sigb,hetMLEdisc$psi), 
                        burnin=nburnin, mcmc=nMCMC,V=Vrw2,thin=Thin)


dfcalhet = as.data.frame(calhet2)
names(dfcalhet) = c("u1","u2","s2f","s2b")
dfcalhet[,1:2] = dfcalhet[,1:2] * 900 +100
dfcalhet[,3] = dfcalhet[,3] * Yv
dfcalhet[,4] = dfcalhet[,4] * Yv
het1 = ggplot(dfcalhet,aes(x=u1,stat(density))) + xlab(expression(K[x]))+ geom_density() + theme_bw() + xlim(50,1000) + geom_vline(aes(xintercept=700),color="red")
het2 = ggplot(dfcalhet,aes(x=u2,stat(density)))+ xlab(expression(K[y])) + geom_density() + theme_bw() +xlim(50,1000)  + geom_vline(aes(xintercept=200),color="red")
het3 = ggplot(dfcalhet,aes(x=s2f,stat(density)))+ xlab(expression(sigma[epsilon]^2)) + geom_density() + theme_bw() + xlim(1,7) + geom_vline(aes(xintercept=4),color="red")
het4 = ggplot(dfcalhet,aes(x=s2b,stat(density))) + xlab(expression(sigma[MD]^2)) + geom_density() + theme_bw() + xlim(1,7) 
```


Calibration with seqhetGP

```{r seqhetGP}
seqhetMLEdisc = RangeEstim(Xfield[,1:2],YfieldnoiseN,seqhet,Uprior,vareps)
calseqhet2 <- MCMCmetrop1R(likcalibrationwithdisc, XF=Xfield[,1:2], yF=YfieldnoiseN, GP=seqhet, Sigdisc=NULL
                        ,logfun=TRUE, theta.init=c(0.5,0.5,1,seqhetMLEdisc$sigb,seqhetMLEdisc$psi), 
                        burnin=nburnin, mcmc=nMCMC,V=Vrw2,thin=Thin)
dfcalseqhet = as.data.frame(calseqhet2)
names(dfcalseqhet) = c("u1","u2","s2f","s2b")
dfcalseqhet[,1:2] = dfcalseqhet[,1:2] * 900 +100
dfcalseqhet[,3] = dfcalseqhet[,3] * Yv
dfcalseqhet[,4] = dfcalseqhet[,4] * Yv
seqhet1 = ggplot(dfcalseqhet,aes(x=u1,stat(density))) + xlab(expression(K[x]))+ geom_density() + theme_bw() + xlim(50,1000) + geom_vline(aes(xintercept=700),color="red")
seqhet2 = ggplot(dfcalseqhet,aes(x=u2,stat(density)))+ xlab(expression(K[y])) + geom_density() + theme_bw() +xlim(50,1000)  + geom_vline(aes(xintercept=200),color="red")
seqhet3 = ggplot(dfcalseqhet,aes(x=s2f,stat(density))) + xlab(expression(sigma[epsilon]^2))+ geom_density() + theme_bw() + xlim(1,7) + geom_vline(aes(xintercept=4),color="red")
seqhet4 = ggplot(dfcalseqhet,aes(x=s2b,stat(density)))+ xlab(expression(sigma[MD]^2))  + geom_density() + theme_bw() + xlim(1,7) 
```


```{r plotcalib}
grid.arrange(arrangeGrob(hom1,hom2,hom3,hom4,top="homGP",ncol=4),arrangeGrob(het1,het2,het3,het4,top="hetGP",ncol=4),
             arrangeGrob(seqhet1,seqhet2,seqhet3,seqhet4,top="seq hetGP",ncol=4),nrow=3,ncol=1)

```


## Prediction

Loading test data 
```{r functionforpred}
test = read.csv("testdata2D.csv",sep=" ")
testdesign = as.matrix(test[,1:2])
Ztest.mean = test[,3]
```


We run prediction from a uniform prior on the calibration parameter and prediction from the posterior distribution of the calibration parameter and the discrepancy.


Note that the seed is not fixed anymore since we use the parallel computations.
```{r runningcalibrationhom}
library(parallel)
npred = nMCMC/Thin
ZpredhomGPnoncal = mclapply(1:npred,prednoncal,GP=hom,vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredhomGPcal = mclapply(1:npred,predcal,GP=hom,cal=calHom2,vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredhomGPnoncal = Reduce(rbind,ZpredhomGPnoncal)
ZpredhomGPcal = Reduce(rbind,ZpredhomGPcal)
```


```{r runningcalibhet}
ZpredhetGPnoncal = mclapply(1:npred,prednoncal,GP=het,vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredhetGPcal = mclapply(1:npred,predcal,GP=het,cal=calhet2,vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredhetGPnoncal = Reduce(rbind,ZpredhetGPnoncal)
ZpredhetGPcal = Reduce(rbind,ZpredhetGPcal)
```


```{r runningcalibseqhet}
ZpredseqhetGPnoncal = mclapply(1:npred,prednoncal,GP=seqhet,vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredseqhetGPcal = mclapply(1:npred,predcal,GP=seqhet,cal=calseqhet2,vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredseqhetGPnoncal = Reduce(rbind,ZpredseqhetGPnoncal)
ZpredseqhetGPcal = Reduce(rbind,ZpredseqhetGPcal)
```


### Some plots

```{r plotspredic}
library(reshape2)
nlocplot=10
head(ZpredhomGPcal)
for (k in 1:nlocplot){
  dfpredhom = data.frame(cal=ZpredhomGPcal[,k],noncal = ZpredhomGPnoncal[,k] )
  df2predhom = melt(dfpredhom)
  ghom = ggplot(df2predhom,aes(x=value, fill=variable)) + geom_density(alpha=0.25)+ theme_bw()  + geom_vline(aes(xintercept=Ztest.mean[k]),color="red")+ggtitle("homGP") 
  dfpredhet = data.frame(cal=ZpredhetGPcal[,k],noncal = ZpredhetGPnoncal[,k] )
  df2predhet = melt(dfpredhet)
  ghet = ggplot(df2predhet,aes(x=value, fill=variable)) + geom_density(alpha=0.25)+ theme_bw()  + geom_vline(aes(xintercept=Ztest.mean[k]),color="red")+ggtitle("hetGP") 
  dfpredseqhet = data.frame(cal=ZpredseqhetGPcal[,k],noncal = ZpredseqhetGPnoncal[,k] )
  df2predsehet = melt(dfpredseqhet)
  ghetseq = ggplot(df2predsehet,aes(x=value, fill=variable)) + geom_density(alpha=0.25)+ theme_bw()  + geom_vline(aes(xintercept=Ztest.mean[k]),color="red")+ggtitle("seqhetGP") 
  print(grid.arrange(ghom,ghet,ghetseq))
}

```


### MSEs and scores



```{r mse}
MSEhomnoncal = mean((rowMeans(ZpredhomGPnoncal)-Ztest.mean)^2)
MSEhomcal = mean((rowMeans(ZpredhomGPcal)-Ztest.mean)^2)
MSEhetnoncal = mean((rowMeans(ZpredhetGPnoncal)-Ztest.mean)^2)
MSEhetcal = mean((rowMeans(ZpredhetGPcal)-Ztest.mean)^2)
MSEseqhetnoncal = mean((rowMeans(ZpredseqhetGPnoncal)-Ztest.mean)^2)
MSEseqhetcal = mean((rowMeans(ZpredseqhetGPcal)-Ztest.mean)^2)
```

```{r scores}
Scorehomnoncal = scoreEstDens(ZpredhomGPnoncal,Ztest.mean)
Scorehomcal = scoreEstDens(ZpredhomGPcal,Ztest.mean)
Scorehetnoncal = scoreEstDens(ZpredhetGPnoncal,Ztest.mean)
Scorehetcal = scoreEstDens(ZpredhetGPcal,Ztest.mean)
Scoreseqhetnoncal = scoreEstDens(ZpredseqhetGPnoncal,Ztest.mean)
Scoreseqhetcal = scoreEstDens(ZpredseqhetGPcal,Ztest.mean)
```




## Other pre-calibration


We try other three other calibrations from L2 minimization, and two guesses (a good and a bad one) for the calibration parameter.


SS minimizer for homGP
```{r L2hom}
L2minhom = optim(c(.5,.5),SumOfSquares,lower=0,upper=1,GP=hom,XF=Xfield[,1:2],yF=YfieldnoiseN)
L2minhom$par* 900 +100
L2minhom$value
SumOfSquares((c(700,200)-100)/900,hom,Xfield[,1:2],YfieldnoiseN)
```

SS minimizer for hetGP
```{r L2het}
L2minhet = optim(c(.5,.5),SumOfSquares,lower=0,upper=1,GP=het,XF=Xfield[,1:2],yF=YfieldnoiseN)
L2minhet$par* 900 +100
L2minhet$value
SumOfSquares((c(700,200)-100)/900,het,Xfield[,1:2],YfieldnoiseN)
```


SS minimizer for seqhetGP
```{r L2seqhet}
L2minseqhet = optim(c(.5,.5),SumOfSquares,lower=0,upper=1,GP=seqhet,XF=Xfield[,1:2],yF=YfieldnoiseN)
L2minseqhet$par* 900 +100
L2minseqhet$value
SumOfSquares((c(700,200)-100)/900,seqhet,Xfield[,1:2],YfieldnoiseN)
```





Prediction for homGP
```{r predhomfixed}
library(parallel)
ZpredhomGPnoncalL2 = mclapply(1:npred,prednoncalfixed,GP=hom,u=L2minhom$par,mc.cores = 8)
ZpredhomGPnoncalL2 = Reduce(rbind,ZpredhomGPnoncalL2)
ZpredhomGPnoncalgoodguess = mclapply(1:npred,prednoncalfixed,GP=hom,u=c(750, 185),vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredhomGPnoncalgoodguess = Reduce(rbind,ZpredhomGPnoncalgoodguess)
ZpredhomGPnoncalbadguess = mclapply(1:npred,prednoncalfixed,GP=hom,u=c(500,260),vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredhomGPnoncalbadguess = Reduce(rbind,ZpredhomGPnoncalbadguess)
```



Prediction for hetGP
```{r predhetfixed}
library(parallel)
ZpredhetGPnoncalL2 = mclapply(1:npred,prednoncalfixed,GP=het,u=L2minhet$par,mc.cores = 8)
ZpredhetGPnoncalL2 = Reduce(rbind,ZpredhetGPnoncalL2)
ZpredhetGPnoncalgoodguess = mclapply(1:npred,prednoncalfixed,GP=het,u=c(750, 185),vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredhetGPnoncalgoodguess = Reduce(rbind,ZpredhetGPnoncalgoodguess)
ZpredhetGPnoncalbadguess = mclapply(1:npred,prednoncalfixed,GP=het,u=c(500,260),vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredhetGPnoncalbadguess = Reduce(rbind,ZpredhetGPnoncalbadguess)
```



Prediction for seqhetGP
```{r predseqhetfixed}
library(parallel)
ZpredseqhetGPnoncalL2 = mclapply(1:npred,prednoncalfixed,GP=seqhet,u=L2minseqhet$par,mc.cores = 8)
ZpredseqhetGPnoncalL2 = Reduce(rbind,ZpredseqhetGPnoncalL2)
ZpredseqhetGPnoncalgoodguess = mclapply(1:npred,prednoncalfixed,GP=seqhet,u=c(750, 185),vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredseqhetGPnoncalgoodguess = Reduce(rbind,ZpredseqhetGPnoncalgoodguess)
ZpredseqhetGPnoncalbadguess = mclapply(1:npred,prednoncalfixed,GP=seqhet,u=c(500,260),vareps=vareps,loc=testdesign,mc.cores = 8)
ZpredseqhetGPnoncalbadguess = Reduce(rbind,ZpredseqhetGPnoncalbadguess)
```


```{r RMSEprecalib}
MSEhomnoncalL2 = mean((rowMeans(ZpredhomGPnoncalL2)-Ztest.mean)^2)
MSEhomnoncalgoodguess = mean((rowMeans(ZpredhomGPnoncalgoodguess)-Ztest.mean)^2)
MSEhomnoncalbadguess = mean((rowMeans(ZpredhomGPnoncalbadguess)-Ztest.mean)^2)
MSEhetnoncalL2 = mean((rowMeans(ZpredhetGPnoncalL2)-Ztest.mean)^2)
MSEhetnoncalgoodguess = mean((rowMeans(ZpredhetGPnoncalgoodguess)-Ztest.mean)^2)
MSEhetnoncalbadguess = mean((rowMeans(ZpredhetGPnoncalbadguess)-Ztest.mean)^2)
MSEseqhetnoncalL2 = mean((rowMeans(ZpredseqhetGPnoncalL2)-Ztest.mean)^2)
MSEseqhetnoncalgoodguess = mean((rowMeans(ZpredseqhetGPnoncalgoodguess)-Ztest.mean)^2)
MSEseqhetnoncalbadguess = mean((rowMeans(ZpredseqhetGPnoncalbadguess)-Ztest.mean)^2)
```


1st row for homGP, 2nd row for hetGP, 3rd row for seqhetGP
1st column for L2 estimate, 2nd column for bad guess, 3rd column for good guess, 4th for unif prior, 5th post calibration
```{r displayfullRMSE}
rbind(c(MSEhomnoncalL2,MSEhomnoncalbadguess,MSEhomnoncalgoodguess,MSEhomnoncal,MSEhomcal),
      c(MSEhetnoncalL2,MSEhetnoncalbadguess,MSEhetnoncalgoodguess,MSEhetnoncal,MSEhetcal),
      c(MSEseqhetnoncalL2,MSEseqhetnoncalbadguess,MSEseqhetnoncalgoodguess,MSEseqhetnoncal,MSEseqhetcal))
```



```{r Scores}
ScorehomnoncalL2 = scoreEstDens(ZpredhomGPnoncalL2,Ztest.mean)
Scorehomnoncalbadguess = scoreEstDens(ZpredhomGPnoncalbadguess,Ztest.mean)
Scorehomnoncalgoodguess = scoreEstDens(ZpredhomGPnoncalgoodguess,Ztest.mean)
ScorehetnoncalL2 = scoreEstDens(ZpredhetGPnoncalL2,Ztest.mean)
Scorehetnoncalbadguess = scoreEstDens(ZpredhetGPnoncalbadguess,Ztest.mean)
Scorehetnoncalgoodguess = scoreEstDens(ZpredhetGPnoncalgoodguess,Ztest.mean)
ScoreseqhetnoncalL2 = scoreEstDens(ZpredseqhetGPnoncalL2,Ztest.mean)
Scoreseqhetnoncalbadguess = scoreEstDens(ZpredseqhetGPnoncalbadguess,Ztest.mean)
Scoreseqhetnoncalgoodguess = scoreEstDens(ZpredseqhetGPnoncalgoodguess,Ztest.mean)
```


1st row for homGP, 2nd row for hetGP, 3rd row for seqhetGP
1st column for L2 estimate, 2nd column for bad guess, 3rd column for good guess, 4th for unif prior, 5th post calibration
```{r displayfullScores}
rbind(c(ScorehomnoncalL2,Scorehomnoncalbadguess,Scorehomnoncalgoodguess,Scorehomnoncal,Scorehomcal),
      c(ScorehetnoncalL2,Scorehetnoncalbadguess,Scorehetnoncalgoodguess,Scorehetnoncal,Scorehetcal),
      c(ScoreseqhetnoncalL2,Scoreseqhetnoncalbadguess,Scoreseqhetnoncalgoodguess,Scoreseqhetnoncal,Scoreseqhetcal))
```



