---
title: "Calibration of the Ocean Simulator"
author: "Pierre Barbillon"
date: "July 1, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
```

```{r packages,message=FALSE}
rm(list=ls())
library(ggplot2)
library(colorRamps)
library(gridExtra)
library(DiceKriging)
library(DiceDesign)
library(hetGP)
library(MCMCpack)
library(lhs)
library(mvtnorm)
library(parallel)
library(reshape2)
```


Loading simulator and GPEs (homGP, hetGP and seqhetGP which is hetGP with a sequential design)
```{r simandGPEs}
# functions
source("SimulatorAndFunctions.R")
source("FunctionCalibration.R")
```

Fixing seed
```{r seed}
seed=1234
set.seed(seed)
```



## Retrieving field data

```{r field}
Field = read.table("FieldData.csv",sep=" ")
Xfield = as.matrix(Field[,1:4])
Yfieldnoise = Field[,5]
n = nrow(Xfield)
Ym <- mean(Yfieldnoise)
Yv <- var(Yfieldnoise)
YfieldnoiseN <- (Yfieldnoise-Ym)/sqrt(Yv)
vareps = 2^2 
```

Simulate discrepancy field
```{r disc}
test = read.csv("testdata2D.csv",sep=" ")
testdesign = as.matrix(test[,1:2])
# simulate discrepancy
Xfielduniloc = unique.matrix(Xfield)
# simulate discrepancy and add it to field data
#disc = simdiscrepancy(rbind(Xfielduniloc[,1:2],testdesign),psi=c(1,2))
#YfieldnoiseN <- (Yfieldnoise-Ym)/sqrt(Yv) + rep(disc[1:nrow(Xfielduniloc)],2)
# or reload it
disc = read.table("discrepancy.csv")[,1]
YfieldnoiseN = read.table("noisyfielddata.csv")[,1]
```


## Emulators

Either
```{r emulator load}
#load("EmulatorCalib4D.Rdata")
```

Or
```{r rebuilding emulator}
## From Static design
Static = read.table("staticDOEcalibration.csv")
Xsim = as.matrix(Static[,1:4])
Ysim = Static[,5]
YsimN = (Ysim-Ym)/sqrt(Yv)  
  
covtype <- "Matern5_2"
noiseControl <- list(g_min=1e-6, g_bounds=c(1e-6, 1), lowerDelta=log(1e-6))
settings <- list(linkThetas="none", initStrategy="smoothed", return.hom=TRUE)
lower <- c(0.01, 0.01, 0.001, 0.001) 
upper <- c(30, 30, 100, 100) 

het <- mleHetGP(Xsim, YsimN, lower=lower, upper=upper, covtype=covtype, noiseControl=noiseControl, 
                settings=settings, maxit=10000)
hom <- het$modHom

## Sequential design
Seq = read.table("seqDOEcalibration.csv")
Xseq = as.matrix(Seq[,1:4])
Yseq = Seq[,5]
YseqN = (Yseq-Ym)/sqrt(Yv)  
  
covtype <- "Matern5_2"
noiseControl <- list(g_min=1e-6, g_bounds=c(1e-6, 1), lowerDelta=log(1e-6))
settings <- list(linkThetas="none", initStrategy="smoothed", return.hom=TRUE)
lower <- c(0.01, 0.01, 0.001, 0.001) 
upper <- c(30, 30, 100, 100) 

seqhet <- mleHetGP(Xseq, YseqN, lower=lower, upper=upper, covtype=covtype, noiseControl=noiseControl, 
                settings=settings, maxit=10000)
```



## Calibration



Settings for MCMC
```{r MCMCsettings}
nMCMC = 1e5  #1e6
nburnin =  2e4 
# variance for random walks
Vrw <- diag(c(.05,.05,.1,.1))/2
Uprior =c(.5,.5)
Thin = 100
```

Calibration with homGP
```{r homGPcalib}
homMLEdisc = RangeEstim(Xfield[,1:2],YfieldnoiseN,hom,Uprior,vareps,rep=2)
# Run MH algo for calibration is time consuming
# calHom <- MCMCmetrop1R(postcalibrationwithdiscrep, XF=Xfield[,1:2], yF=YfieldnoiseN, GP=hom,
#                        Sigdisc=homMLEdisc$Sigdisc,priorUpBounds2b=3/Yv,rep=2,logvar=T
#                              ,logfun=TRUE, theta.init=c(0.5,0.5,log(2/Yv),log(2/Yv)),
#                              burnin=nburnin, mcmc=nMCMC,V=Vrw,thin=Thin)
# write.table(calHom,file="calHom.csv")


# reloading saved calibration results
calHom = read.table("calHom.csv",h=T,sep=" ")
calHom = as.matrix(calHom)
dfcalhom = as.data.frame(calHom)
names(dfcalhom) = c("u1","u2","s2f","s2b")
dfcalhom[,1:2] = dfcalhom[,1:2] * 900 +100
dfcalhom[,3] = exp(dfcalhom[,3]) * Yv
dfcalhom[,4] = exp(dfcalhom[,4]) * Yv
hom1 = ggplot(dfcalhom,aes(x=u1,stat(density)))+ xlab(expression(K[x])) + geom_density() + theme_bw() + xlim(50,1000) + geom_vline(aes(xintercept=700),color="red")
hom2 = ggplot(dfcalhom,aes(x=u2,stat(density))) + xlab(expression(K[y])) + geom_density() + theme_bw() +xlim(50,1000)  + geom_vline(aes(xintercept=200),color="red")
hom3 = ggplot(dfcalhom,aes(x=s2f,stat(density)))+ xlab(expression(sigma[epsilon]^2))  + geom_density() + theme_bw() + xlim(0,7) + geom_vline(aes(xintercept=4),color="red")
hom4 = ggplot(dfcalhom,aes(x=s2b,stat(density)))+ xlab(expression(sigma[MD]^2)) + geom_density() + theme_bw() + xlim(0,3) +geom_vline(aes(xintercept=Yv*.02),color="red")
```



Calibration with hetGP
```{r hetGPcalib}
hetMLEdisc = RangeEstim(Xfield[,1:2],YfieldnoiseN,het,Uprior,vareps,rep=2)
# Run MH algo for calibration is time consuming
# calHet <- MCMCmetrop1R(postcalibrationwithdiscrep, XF=Xfield[,1:2], yF=YfieldnoiseN, GP=het,
#                              Sigdisc=hetMLEdisc$Sigdisc,priorUpBounds2b=3/Yv,rep=2,logvar=T
#                              ,logfun=TRUE, theta.init=c(0.5,0.5,log(2/Yv),log(2/Yv)),
#                              burnin=nburnin, mcmc=nMCMC,V=Vrw,thin=Thin)
# write.table(calHet,file="calHet.csv")

# reloading saved calibration results
calHet = read.table("calHet.csv",h=T,sep=" ")
calHet = as.matrix(calHet)
dfcalhet = as.data.frame(calHet)
names(dfcalhet) = c("u1","u2","s2f","s2b")
dfcalhet[,1:2] = dfcalhet[,1:2] * 900 +100
dfcalhet[,3] = exp(dfcalhet[,3]) * Yv
dfcalhet[,4] = exp(dfcalhet[,4]) * Yv
het1 = ggplot(dfcalhet,aes(x=u1,stat(density))) + xlab(expression(K[x]))+ geom_density() + theme_bw() + xlim(50,1000) + geom_vline(aes(xintercept=700),color="red")
het2 = ggplot(dfcalhet,aes(x=u2,stat(density)))+ xlab(expression(K[y])) + geom_density() + theme_bw() +xlim(50,1000)  + geom_vline(aes(xintercept=200),color="red")
het3 = ggplot(dfcalhet,aes(x=s2f,stat(density)))+ xlab(expression(sigma[epsilon]^2)) + geom_density() + theme_bw() + xlim(0,7) + geom_vline(aes(xintercept=4),color="red")
het4 = ggplot(dfcalhet,aes(x=s2b,stat(density))) + xlab(expression(sigma[MD]^2)) + geom_density() + theme_bw() + xlim(0,3) +geom_vline(aes(xintercept=Yv*.02),color="red")
```


Calibration with seqhetGP

```{r seqhetGP}
seqhetMLEdisc = RangeEstim(Xfield[,1:2],YfieldnoiseN,seqhet,Uprior,vareps,rep=2)
# Run MH algo for calibration is time consuming
# calSeqHet <- MCMCmetrop1R(postcalibrationwithdiscrep, XF=Xfield[,1:2], yF=YfieldnoiseN, GP=seqhet,
#                                 Sigdisc=seqhetMLEdisc$Sigdisc,priorUpBounds2b=3/Yv,rep=2,logvar=T
#                              ,logfun=TRUE, theta.init=c(0.5,0.5,log(2/Yv),log(2/Yv)),
#                              burnin=nburnin, mcmc=nMCMC,V=Vrw,thin=Thin)
#write.table(calSeqHet,file="calSeqHet.csv")

# reloading saved calibration results
calSeqHet = read.table("calSeqHet.csv",h=T,sep=" ")
calSeqHet = as.matrix(calSeqHet)
dfcalseqhet = as.data.frame(calSeqHet)
names(dfcalseqhet) = c("u1","u2","s2f","s2b")
dfcalseqhet[,1:2] = dfcalseqhet[,1:2] * 900 +100
dfcalseqhet[,3] = exp(dfcalseqhet[,3]) * Yv
dfcalseqhet[,4] = exp(dfcalseqhet[,4]) * Yv
seqhet1 = ggplot(dfcalseqhet,aes(x=u1,stat(density))) + xlab(expression(K[x]))+ geom_density() + theme_bw() + xlim(50,1000) + geom_vline(aes(xintercept=700),color="red")
seqhet2 = ggplot(dfcalseqhet,aes(x=u2,stat(density)))+ xlab(expression(K[y])) + geom_density() + theme_bw() +xlim(50,1000)  + geom_vline(aes(xintercept=200),color="red")
seqhet3 = ggplot(dfcalseqhet,aes(x=s2f,stat(density))) + xlab(expression(sigma[epsilon]^2))+ geom_density() + theme_bw() + xlim(0,7) + geom_vline(aes(xintercept=4),color="red")
seqhet4 = ggplot(dfcalseqhet,aes(x=s2b,stat(density)))+ xlab(expression(sigma[MD]^2))  + geom_density() + theme_bw() + xlim(0,3) +geom_vline(aes(xintercept=Yv*.02),color="red")
```


```{r plotcalib}
grid.arrange(arrangeGrob(hom1,hom2,hom3,hom4,top="homGP",ncol=4),arrangeGrob(het1,het2,het3,het4,top="hetGP",ncol=4),
             arrangeGrob(seqhet1,seqhet2,seqhet3,seqhet4,top="seq hetGP",ncol=4),nrow=3,ncol=1)

```


## Prediction

Loading test data 
```{r functionforpred}
# test = read.csv("testdata2D.csv",sep=" ")
# testdesign = as.matrix(test[,1:2])
Ztest.mean = test[,3] + sqrt(Yv)*disc[-c(1:nrow(Xfielduniloc))]
Ztest.sim = Ztest.mean + rnorm(length(Ztest.mean),0,sqrt(vareps))
```


We run prediction from a uniform prior on the calibration parameter and prediction from the posterior distribution of the calibration parameter and the discrepancy.


Note that the seed is not fixed anymore since we use the parallel computations.
```{r runningcalibrationhom}
npred = nMCMC/Thin
ZpredhomGPnoncal = mclapply(1:npred,prednoncal,GP=hom,vareps=vareps,loc=testdesign,Ym=Ym,Yv=Yv,mc.cores = 8)
ZpredhomGPcal = mclapply(1:npred,predcal,GP=hom,cal=calHom,vareps=vareps,
                         loc=testdesign,YfN=YfieldnoiseN,Xfield=Xfield,Ym=Ym,Yv=Yv,psi=homMLEdisc$psi
                         ,mc.cores = 8)
ZpredhomGPnoncal = Reduce(rbind,ZpredhomGPnoncal)
ZpredhomGPcal = Reduce(rbind,ZpredhomGPcal)
```


```{r runningcalibhet}
ZpredhetGPnoncal = mclapply(1:npred,prednoncal,GP=het,vareps=vareps,loc=testdesign,Ym=Ym,Yv=Yv,mc.cores = 8)
ZpredhetGPcal = mclapply(1:npred,predcal,GP=het,cal=calHet,
                         vareps=vareps,loc=testdesign,YfN=YfieldnoiseN,Xfield=Xfield,Ym=Ym,Yv=Yv,psi=hetMLEdisc$psi
                         ,mc.cores = 8)
ZpredhetGPnoncal = Reduce(rbind,ZpredhetGPnoncal)
ZpredhetGPcal = Reduce(rbind,ZpredhetGPcal)
```


```{r runningcalibseqhet}
ZpredseqhetGPnoncal = mclapply(1:npred,prednoncal,GP=seqhet,vareps=vareps,loc=testdesign,Ym=Ym,Yv=Yv,mc.cores = 8)
ZpredseqhetGPcal = mclapply(1:npred,predcal,GP=seqhet,cal=calSeqHet,
                            vareps=vareps,loc=testdesign,YfN=YfieldnoiseN,Xfield=Xfield,Ym=Ym,Yv=Yv,psi=seqhetMLEdisc$psi,
                            mc.cores = 8)
ZpredseqhetGPnoncal = Reduce(rbind,ZpredseqhetGPnoncal)
ZpredseqhetGPcal = Reduce(rbind,ZpredseqhetGPcal)
```


### Some plots

```{r plotspredic}
nlocplot=10
for (k in 1:nlocplot){
    dfpredhom = data.frame(cal=ZpredhomGPcal[,k],noncal = ZpredhomGPnoncal[,k] )
  df2predhom = melt(dfpredhom)
  ghom = ggplot(df2predhom,aes(x=value, fill=variable)) + geom_density(alpha=0.25)+ theme_bw()  + geom_vline(aes(xintercept=Ztest.mean[k]),color="red")+ggtitle("homGP")
  dfpredhet = data.frame(cal=ZpredhetGPcal[,k],noncal = ZpredhetGPnoncal[,k] )
  df2predhet = melt(dfpredhet)
  ghet = ggplot(df2predhet,aes(x=value, fill=variable)) + geom_density(alpha=0.25)+ theme_bw()  + geom_vline(aes(xintercept=Ztest.mean[k]),color="red")+ggtitle("hetGP")
  dfpredseqhet = data.frame(cal=ZpredseqhetGPcal[,k],noncal = ZpredseqhetGPnoncal[,k] )
  df2predsehet = melt(dfpredseqhet)
  ghetseq = ggplot(df2predsehet,aes(x=value, fill=variable)) + geom_density(alpha=0.25)+ theme_bw()  + geom_vline(aes(xintercept=Ztest.mean[k]),color="red")+ggtitle("seqhetGP")
  print(grid.arrange(ghom,ghet,ghetseq))
}
```


### MSEs and scores



```{r mse}
RMSEhomnoncal = sqrt(mean((rowMeans(ZpredhomGPnoncal)-Ztest.mean)^2))
RMSEhomcal = sqrt(mean((rowMeans(ZpredhomGPcal)-Ztest.mean)^2))
RMSEhetnoncal = sqrt(mean((rowMeans(ZpredhetGPnoncal)-Ztest.mean)^2))
RMSEhetcal = sqrt(mean((rowMeans(ZpredhetGPcal)-Ztest.mean)^2))
RMSEseqhetnoncal = sqrt(mean((rowMeans(ZpredseqhetGPnoncal)-Ztest.mean)^2))
RMSEseqhetcal = sqrt(mean((rowMeans(ZpredseqhetGPcal)-Ztest.mean)^2))
```

```{r scores}
Scorehomnoncal = scoreEstDens(ZpredhomGPnoncal,Ztest.sim)
Scorehomcal = scoreEstDens(ZpredhomGPcal,Ztest.sim)
Scorehetnoncal = scoreEstDens(ZpredhetGPnoncal,Ztest.sim)
Scorehetcal = scoreEstDens(ZpredhetGPcal,Ztest.sim)
Scoreseqhetnoncal = scoreEstDens(ZpredseqhetGPnoncal,Ztest.sim)
Scoreseqhetcal = scoreEstDens(ZpredseqhetGPcal,Ztest.sim)
```




## Other pre-calibration


We try other three other calibrations from L2 minimization, and a guess for the calibration parameter.


SS minimizer for homGP
```{r L2hom}
L2minhom = optim(c(.5,.5),SumOfSquares,lower=0,upper=1,GP=hom,XF=Xfield[,1:2],yF=YfieldnoiseN)
L2minhom$par* 900 +100
L2minhom$value
SumOfSquares((c(700,200)-100)/900,hom,Xfield[,1:2],YfieldnoiseN)
```

SS minimizer for hetGP
```{r L2het}
L2minhet = optim(c(.5,.5),SumOfSquares,lower=0,upper=1,GP=het,XF=Xfield[,1:2],yF=YfieldnoiseN)
L2minhet$par* 900 +100
L2minhet$value
SumOfSquares((c(700,200)-100)/900,het,Xfield[,1:2],YfieldnoiseN)
```


SS minimizer for seqhetGP
```{r L2seqhet}
L2minseqhet = optim(c(.5,.5),SumOfSquares,lower=0,upper=1,GP=seqhet,XF=Xfield[,1:2],yF=YfieldnoiseN)
L2minseqhet$par* 900 +100
L2minseqhet$value
SumOfSquares((c(700,200)-100)/900,seqhet,Xfield[,1:2],YfieldnoiseN)
```


We propose a guess. 

```{r goodbadguesses}
guess = (c(600, 300) -100)/900 
```



Prediction for homGP
```{r predhomfixed}
ZpredhomGPnoncalL2 = mclapply(1:npred,prednoncalfixed,GP=hom,u=L2minhom$par,vareps=vareps,Ym=Ym,Yv=Yv,loc=testdesign,mc.cores = 8)
ZpredhomGPnoncalL2 = Reduce(rbind,ZpredhomGPnoncalL2)
ZpredhomGPnoncalguess = mclapply(1:npred,prednoncalfixed,GP=hom,u=guess,vareps=vareps,Ym=Ym,Yv=Yv,loc=testdesign,mc.cores = 8)
ZpredhomGPnoncalguess = Reduce(rbind,ZpredhomGPnoncalguess)
```



Prediction for hetGP
```{r predhetfixed}
ZpredhetGPnoncalL2 = mclapply(1:npred,prednoncalfixed,GP=het,u=L2minhet$par,vareps=vareps,Ym=Ym,Yv=Yv,loc=testdesign,mc.cores = 8)
ZpredhetGPnoncalL2 = Reduce(rbind,ZpredhetGPnoncalL2)
ZpredhetGPnoncalguess = mclapply(1:npred,prednoncalfixed,GP=het,u=guess,vareps=vareps,Ym=Ym,Yv=Yv,loc=testdesign,mc.cores = 8)
ZpredhetGPnoncalguess = Reduce(rbind,ZpredhetGPnoncalguess)
```



Prediction for seqhetGP
```{r predseqhetfixed}
ZpredseqhetGPnoncalL2 = mclapply(1:npred,prednoncalfixed,GP=seqhet,u=L2minseqhet$par,vareps=vareps,Ym=Ym,Yv=Yv,loc=testdesign,mc.cores = 8)
ZpredseqhetGPnoncalL2 = Reduce(rbind,ZpredseqhetGPnoncalL2)
ZpredseqhetGPnoncalguess = mclapply(1:npred,prednoncalfixed,GP=seqhet,u=guess,vareps=vareps,Ym=Ym,Yv=Yv,loc=testdesign,mc.cores = 8)
ZpredseqhetGPnoncalguess = Reduce(rbind,ZpredseqhetGPnoncalguess)
```




```{r RMSEprecalib}
RMSEhomnoncalL2 = sqrt(mean((rowMeans(ZpredhomGPnoncalL2)-Ztest.mean)^2))
RMSEhomnoncalguess = sqrt(mean((rowMeans(ZpredhomGPnoncalguess)-Ztest.mean)^2))
RMSEhetnoncalL2 = sqrt(mean((rowMeans(ZpredhetGPnoncalL2)-Ztest.mean)^2))
RMSEhetnoncalguess = sqrt(mean((rowMeans(ZpredhetGPnoncalguess)-Ztest.mean)^2))
RMSEseqhetnoncalL2 = sqrt(mean((rowMeans(ZpredseqhetGPnoncalL2)-Ztest.mean)^2))
RMSEseqhetnoncalguess = sqrt(mean((rowMeans(ZpredseqhetGPnoncalguess)-Ztest.mean)^2))
```



1st row for homGP, 2nd row for hetGP, 3rd row for seqhetGP
1st column for L2 estimate, 2nd column for guess, 3rd column for unif prior, 4th post calibration
```{r displayfullRMSE}
rbind(c(RMSEhomnoncalL2,RMSEhomnoncalguess,RMSEhomnoncal,RMSEhomcal),
      c(RMSEhetnoncalL2,RMSEhetnoncalguess,RMSEhetnoncal,RMSEhetcal),
      c(RMSEseqhetnoncalL2,RMSEseqhetnoncalguess,RMSEseqhetnoncal,RMSEseqhetcal))
```



```{r Scores}
ScorehomnoncalL2 = scoreEstDens(ZpredhomGPnoncalL2,Ztest.sim)
Scorehomnoncalguess = scoreEstDens(ZpredhomGPnoncalguess,Ztest.sim)
ScorehetnoncalL2 = scoreEstDens(ZpredhetGPnoncalL2,Ztest.sim)
Scorehetnoncalguess = scoreEstDens(ZpredhetGPnoncalguess,Ztest.sim)
ScoreseqhetnoncalL2 = scoreEstDens(ZpredseqhetGPnoncalL2,Ztest.sim)
Scoreseqhetnoncalguess = scoreEstDens(ZpredseqhetGPnoncalguess,Ztest.sim)
```


1st row for homGP, 2nd row for hetGP, 3rd row for seqhetGP
1st column for L2 estimate, 2nd column for guess, 3rd column for unif prior, 4th post calibration
```{r displayfullScores}
rbind(c(ScorehomnoncalL2,Scorehomnoncalguess,Scorehomnoncal,Scorehomcal),
      c(ScorehetnoncalL2,Scorehetnoncalguess,Scorehetnoncal,Scorehetcal),
      c(ScoreseqhetnoncalL2,Scoreseqhetnoncalguess,Scoreseqhetnoncal,Scoreseqhetcal))
```
